%!TEX root = ../thesis-summary.tex

\section{Related Work}
\label{section:related-work}

When considering pub-sub systems, there is a set of different options that will
lay ground for the behaviour of the whole system. We call these options, design
dimensions. Specifically, in our case, one of the biggest decisions when
designing a pub-sub system is what kind of subscription model to use. The
subscription model determines how subscribers will define which events they are
interested in. There are three major approaches covered by relevant literature
\cite{Kermarrec2013} \cite{Eugster2003} and that implementations usually
follow:

\begin{itemize}
  \item
    Topic based subscriptions - Clients subscribe to classes of events, usually identified by keywords.\cite{Castro2002}\cite{Zhuang2001}\cite{Baldoni2007}\cite{Apolonia2018}\cite{Setty2012}
  \item
    Content based subscriptions - Clients subscribe to events based on specific values (or ranges of values) on the properties of the events.\cite{Strom1998}\cite{Cugola2001}\cite{Carzaniga2003}\cite{Gupta2004}\cite{Bharambe2002}\cite{Voulgaris2005}
  \item
    Type based subscriptions\cite{Eugster2000} - Bring the notion of a type scheme to a topic based subscription model.\cite{Pietzuch2002}
\end{itemize}

The subscription models are tied to the expressiveness of the system as a
whole. Point in case, a content based subscription model allows for a lot more
expressiveness in subscription definition, but it makes it a lot harder to
implement a scalable way of filtering messages.

Another key design dimension, specially when covering P2P pub-sub systems, is
how peers choose to organise and maintain their view of the underlying network
(commonly referred to as network overlays). 

The systems we will cover next have chosen different approaches for the design
dimensions described above, however all of them have played a seminal role for
our proposed solution.

\textbf{Gryphon} \cite{Strom1998} is a content based pub-sub system built on
top of a centralised broker hierarchy topology, successfully deployed over the
Internet for real-time sports score distribution at the Grand Slam Tennis
events, Ryder Cup, and for monitoring and statistics reporting at the Sydney
Olympics~\footnote{\url{https://www.research.ibm.com/distributedmessaging/gryphon.html}}.
Developed at IBM, Gryphon uses an interesting approach to match events with
subscriptions \cite{Aguilera1999}, relying on a distributed broker based
network to build a tree structure representing the subscription schema. 

\textbf{Siena} \cite{Carzaniga2003} is a content based pub-sub system built on
top of a centralised broker mesh topology. Siena does not make any assumptions
on how the communication between servers and client-server works, as this is
not vital for the system to work. Instead, for server to server communication,
it provides a set of options ranging from P2P communication to a more
hierarchical structure, each with its respective advantages and shortcomings.
Still on the subject of broker based network solutions, HyperPubSub
\cite{Zupan2017} is a recent example of such approach but focused on bringing
verifiability and other forms of decentralised validation to pub-sub
operations.

\textbf{Scribe} \cite{Castro2002} is a topic based pub-sub system built
on top of a fully decentralised network (P2P). In order to do this, it
relies on Pastry DHT as its overlay structure. This allows it to
leverage the robustness, self-organisation, locality and reliability
properties of Pastry to build a set of per-topic multicast trees used to disseminate events.

\textbf{Meghdoot} \cite{Gupta2004} is a content based pub-sub system.  It is
built on top of a P2P network, specifically CAN DHT \cite{Ratnasamy2001a}.
Meghdoot leverages the multidimensional space provided by the CAN DHT in order
to create an expressive content based system.

\textbf{Poldercast} \cite{Setty2012} is a recent pub-sub system with a strong
focus on scalability, robustness, efficiency and fault tolerance. It follows a
topic based model and follows a fully decentralised architecture. The key
detail about this system is that it tries to blend deterministic propagation
over a structured overlay, with probabilistic dissemination through gossip
based unstructured overlays. In order to do this, Poldercast uses 3 different
overlays.\cite{Voulgaris2013}\cite{Voulgaris2005a} Similar to systems like
VCube-PS \cite{DeAraujo2017}, only nodes subscribed to a topic will receive
events published to that topic. In other words, no relay nodes are used. It
also focuses on handling churn through the use of a mixture of gossip
mechanisms, ensuring a highly resilient network.  Finally, it seeks to reduce
message duplication factor (i.e. nodes receiving the same message more than
once).

Architecturally speaking, one can see the similarities between Poldercast and
SELECT \cite{Apolonia2018}, as it too relies on a set of three different
overlays, working together to fulfil subscriptions. However, SELECT brings a
different set of properties to the table, as it maps the social connection
graph of the peers to the actual overlays operating underneath. This way, the
system can exploit both the social graph and the online activity of each social
user (each peer) to establish connections and disseminate messages accordingly
and avoid an unnecessary number of hops.

The aim of our work is also to take advantage of the network infrastructure and technologies already in place. One of the best ways of doing so is by leveraging what the Web platform has to offer.

TODO Mention Javascript, browsserCloud,js, IPFS, libp2p, Merkle DAG
