%!TEX root = ../dissertation.tex

% Acknowledgments: This one is optional
\chapter*{Acknowledgments}

The potential for this list to be a whole other dissertation is enormous, but I
will do my best to keep it short.

Starting by a very special thank you to my dissertation supervisors Luís Veiga
and David Dias, who set me up to this fantastic challenge and helped me out
along the way. To João Tiago, a dear colleague and a friend, who was also
writing his master thesis and provided me with tireless support. To Microsoft
Azure and especially my great friend Jorge Palma, for supporting our test
executions in their cloud infrastructure.

To all my colleagues over at YLD who gave me room and support to keep on going,
especially Filipe Pinheiro (with his premium Grammarly account) and Fábio
Oliveira.

To countless friends and colleagues that either through their kind words or
just by asking how was my dissertation work going, gave me the strength to push
it through.

To my family, my parents and specially to my two brothers, Pedro Beja, who gave
me my first computer and made this whole thing possible and Fábio Antunes who,
besides being the best engineer and the best colleague one could ever wish for,
is the main reason I got into computer science in the first place.

Last but not least, to my girlfriend Rita Cravo. For making sure I kept being a
functioning human being, for never giving up on me, for being there for me on
all those all-nighters, for all the weekends that I had to spend in front of
the computer and for all the plans we had to postpone. Thank you, from the
bottom of my heart. This document would never exist without you.

